---
title: "Linear Models with L2 Regularization"
author: "Chadd Frasier"
date: "6/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.show = 'hold', fig.height = 3)
set.seed(1023)
```

## Linear Models L2 Regularization

This is a Linear model using L2 reguarization that was writing in pure R using gradient descent of the mean squared loss. This model is also known as Ridge Regression. In this report I will use the data found in a single band to predict other value of that same band, while holding out test sets for cross validation.

### Square Loss (Regression): $L(w) = ||Xw + {\beta}_n-y||^2_2$

```{r InitializationStep}
# Import libraries
library(planetLearn)
library(ggplot2)
library(knitr)

# load the data file from the compressed library
load(file = "../../data/lunarData.rda")


limit <- 300000
limiter <- seq_len(limit)

# Isolate the observations and features
X.mat <- mapply(as.numeric, lunarData[limiter,-1])
y.vec <- as.numeric(lunarData[limiter,1])

X.test <- mapply(as.numeric, lunarData[limit:(limit+10000), -1])
y.test <- as.numeric(lunarData[limit:(limit+10000), 1])

n.folds <- 1000
fold.vec <- sample(n.folds, replace = TRUE, size = nrow(X.mat))

# get the penalty vector
penalty.vec <- seq(10, 0.1, by = -0.1)

# plot the data
knitr::kable( head(lunarData) )
```

## Training

```{r TrainingStep}
# run the early stopping algorithm with a set max of iterations
earlystopping.list <- LMSquareLossEarlyStoppingCV(X.mat, y.vec, fold.vec, 500L, step.size=.0000035 )

# run without early stop
#L2.list <- LMSquareLossL2CV(X.mat, y.vec, fold.vec, penalty.vec)

earlystopping.predict <- earlystopping.list$predict(as.matrix(X.test) )
#L2.predict <- L2.list$predict(X.test)
baseline.predict <- mean(y.test)

# L2 loss
earlystopping.loss <- mean((earlystopping.predict - y.test) ^ 2)
#L2.loss <- mean((L2.predict - y.test) ^ 2)
baseline.loss <- mean((baseline.predict - y.test) ^ 2)
```

```{r Plot}
plotDataV <- data.frame(ValidationLoss=earlystopping.list$mean.validation.loss.vec, Iterations=seq_len(length(earlystopping.list$mean.validation.loss.vec)))
plotDataT <- data.frame(TrainLoss=earlystopping.list$mean.train.loss.vec, Iterations=seq_len(length(earlystopping.list$mean.train.loss.vec)))
p2 <- ggplot() +
  geom_line(data = plotDataT, aes(x=Iterations, y=TrainLoss), color="blue") +
  geom_line(data=plotDataV, aes(y=ValidationLoss, x=Iterations), color="red") +
  xlab("Iterations") +
  ylab("Squared Loss") +
  labs(title = "Loss over Iterations") 
plot(p2)
```

## Conclusion

Conclusion will go here